---
title: "Descriptive Report of Lead & IQ Data"
output:
  html_document: default
  pdf_document: default
date: "2024-10-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(readr)

# Read in data
setwd("C:/Week6/Meier_Project_01")
data <- read_csv("DataRaw/lead-iq-01.csv")

# Save data in processed
save(data, file = file.path("DataProcessed", "data.RData"))  # Saves as .RData file
```

## Graph showing IQ levels by location status
Histogram displays distribution of IQ by distance from the smelter.
```{r histogram, echo=TRUE}
ggplot(data, aes(x = IQ, fill = Smelter)) + 
    geom_histogram(position = "identity", alpha = 0.5, bins = 10, color = "black", na.rm = TRUE) + 
    labs(title = "Histogram of IQ Levels by Smelter Status", 
         x = "IQ Levels", 
         y = "Count") + 
    theme_minimal()
```

## Summary table of IQ by Smelter distance
```{r summary_table, echo=FALSE}
# Create a summary table calculating average IQ by Smelter status

summary_table <- data %>%
    group_by(Smelter) %>%
    summarise(
        Count = n(),
        Average_IQ = mean(IQ, na.rm = TRUE),
        Min_IQ = min(IQ, na.rm = TRUE),
        Max_IQ = max(IQ, na.rm = TRUE)
    ) %>%
    arrange(Smelter)  # Optional: arrange by Smelter status

# Format the table using kable
kable(summary_table, 
      caption = "Summary Table of Average IQ by Smelter Location",
      format = "html",   # Change to "latex" for PDF output
      digits = 2,        # Number of decimal places
      col.names = c("Smelter Location", "Count", "Average IQ", "Min IQ", "Max IQ")) %>%  # Custom column names
  
  kable_styling(full_width = FALSE) %>%
  column_spec(1, width = "3cm") %>%  # Adjust column 1 width (Smelter Status)
  column_spec(2, width = "2cm") %>%  # Adjust column 2 width (Count)
  column_spec(3, width = "3cm") %>%  # Adjust column 3 width (Average IQ)
  column_spec(4, width = "2cm") %>%  # Adjust column 4 width (Min IQ)
  column_spec(5, width = "2cm")      # Adjust column 5 width (Max IQ)

```
## Discussion
Table summarizes the count, average, and range of IQ levels for people far and near to the smelter.The average IQ for individuals near the smelter is `r round(summary_table$Average_IQ[summary_table$Smelter == "Near"], 2)`, while for those living farther away, the average IQ is `r round(summary_table$Average_IQ[summary_table$Smelter == "Far"], 2)`.
